---
output:
  word_document: default
  html_document: default
---
##women breast cancer: diagnostic
```{r}
bc_d=read.csv("C:/Temp/data/wisconson/wdbc.data", 
              header=TRUE)

dim(bc_d)

head(bc_d)

colnames(bc_d) = c("ID", "Diagnosis", 
                   "Radius_mean", "Texture_mean", "Parameter_mean", "Area_mean", "Smoothness_mean",
                   "Compactness_mean", "Concavity_mean", "Concave_points_mean", "Symmetry_mean", "Fraction_dimension_mean",
                   "Radius_se", "Texture_se", "Parameter_se", "Area_se", "Smoothness_se",
                   "Compactness_se", "Concavity_se", "Concave_points_se", "Symmetry_se", "Fraction_dimension_se",
                   "Radius_worst", "Texture_worst", "Parameter_worst", "Area_worst", "Smoothness_worst",
                   "Compactness_worst", "Concavity_worst", "Concave_points_worst", "Symmetry_worst", "Fraction_dimension_worst")

head(bc_d)
#diagnosis: M(악성), B(양성)
bc_d$Diagnosis = as.factor(bc_d$Diagnosis)
```

##################################################################
#                 EDA
##################################################################
```{r}
sum(is.na(bc_d))
```
#no missing values

#data filtering
```{r}
str(bc_d)
table(bc_d$Diagnosis)
```

####################################################################
#       PCA
####################################################################
```{r}
bcX = bc_d[,-c(1,2)]
head(bcX)

bcX.sc = as.matrix(scale(bcX))
```


#original class
```{r}
p_col = bc_d$Diagnosis
p_col

head(data.frame(apply(bcX.sc, 2, function(x) (x-mean(x))/sd(x))))
head(bcX.sc)

library(mulgar)
library(ggplot2)
bc_pca = prcomp(bcX.sc)
bc_scree = ggscree(bc_pca, q=5) + theme_minimal()


summary(bc_pca) 
```
#around 3 PCs will cover 72.6%


#lets draw a 3-dim plot & PCA1~PCA3
```{r}
library(dplyr)
library(gt)
library(tourr)
bc_pca$rotation[,1:3] %>%
  as_tibble(rownames="Variable") %>% 
  arrange(desc(PC1), desc(PC2), desc(PC3)) %>%
  gt() %>%
  fmt_number(columns = c(PC1, PC2, PC3),
             decimals = 2)

animate(bc_pca$x[,1:3],
           tour_path = grand_tour(), 
           display=display_xy(cex=1))
```
#a few outliers appear


```{r}
animate_xy(bc_pca$x[,1:3], col=p_col)
```
#the projection seems well done, since we can distinguish the two different diagnosis status.
#It seems that the two diagnosis do seem very close, in some viewpoint they can be separated.
#maybe we can use the LDA/QDA?


##compare with original data
```{r}
animate_xy(bcX.sc, col=p_col, cex=1, 
           tour_path=grand_tour())
```


library(GGally)
```{r}
bc_pca.dat = as.matrix(cbind(bc_pca$x[,1:3], p_col))
bc_pca.dat = transform(bc_pca.dat, PC1 = as.numeric(PC1))
bc_pca.dat = transform(bc_pca.dat, PC2 = as.numeric(PC2))
bc_pca.dat = transform(bc_pca.dat, PC3 = as.numeric(PC3))
bc_pca.dat$p_col = as.factor(bc_pca.dat$p_col)
str(bc_pca.dat)
```


```{r}
library(GGally)
ggscatmat(bc_pca.dat, columns=1:3, color="p_col")
```
#PC1&PC2 has the best clustering
#PC1&PC3 also has good clustering




####################################################################
#       LDA (with PCA results)
####################################################################
```{r}
library(MASS)
table(p_col)  #original data's classification

bc_lda = lda(p_col ~ PC1 + PC2 + PC3 ,data=bc_pca.dat,
             prior=c(1/2, 1/2))
options(digits=2)
bc_lda

bc_lda$scaling

bc_lda_pred = predict(bc_lda, bc_pca.dat)
bc_lda_pred$class  #1:B, 2:M
```


#오분류율(쫌 쎈데..)
#p_col = ifelse(p_col=="B", 1,2)
```{r}
table(ifelse(bc_lda_pred$class==1,"B", "M"), p_col)
mean(ifelse(bc_lda_pred$class==1,"B", "M")!= p_col)
```
#0.053

```{r}
ldahist(bc_lda_pred$x, g=bc_lda_pred$class)

bc_pca.dat$pred =bc_lda_pred$class
bc_pca.dat$miss = bc_lda_pred$class != as.factor(ifelse(bc_d$Diagnosis=="B",1,2))

sum(bc_pca.dat$miss)
```
#30개


##plot
```{r}
ggplot(bc_pca.dat, aes(PC1, PC2))+
  geom_point(data = bc_pca.dat[bc_pca.dat$miss,],
             col="red",size= 4)+  #missing ones are in red dot
  geom_point(aes(color=factor(p_col)))  #the actual(original) classificaiton
```
#1 = B, 2 = M
#the ones in between seem like they have been misclassified
#It seems that more 2(M=악성종양) have been misclassified than 1.
#This could be dangerous, as those who have cancer could be classified as if they are fine.


```{r}
ggplot(bc_pca.dat, aes(PC1, PC3))+
  geom_point(data = bc_pca.dat[bc_pca.dat$miss,],
             col="red",size= 4)+  #missing ones are in red dot
  geom_point(aes(color=factor(p_col)))  #the actual(original) classificaiton
```
#1 = B, 2 = M
#The 구분 is not as precise as the first one.
#There are more 1(B=양성) that have been misclassified as 2
#This might be a safer model for patients

```{r}
ggplot(bc_pca.dat, aes(PC2, PC3))+
  geom_point(data = bc_pca.dat[bc_pca.dat$miss,],
             col="red",size= 4)+  #missing ones are in red dot
  geom_point(aes(color=factor(p_col)))  #the actual(original) classificaiton
```
#1 = B, 2 = M
#Bad PCA


###3d plot
```{r}
animate_xy(bc_pca$x[,1:3], col=bc_lda_pred$class)
```
#compare with original coloring
```{r}
animate_xy(bc_pca$x[,1:3], col=p_col)
```



####################################################################
#       LDA (with original data)
####################################################################
```{r}
table(p_col)  #original data's classification

bc_lda_org = lda(Diagnosis~., data=bc_d[,-1],
                 prior=c(1/2, 1/2))
  
options(digits=2)
bc_lda_org

bc_lda_org$scaling

bc_lda_org.pred = predict(bc_lda_org, bc_d[,-1])
bc_lda_org.pred$class  #1:B, 2:M
```

#오분류율
```{r}
table(bc_lda_org.pred$class, bc_d$Diagnosis)
mean(bc_lda_org.pred$class !=bc_d$Diagnosis)
```
#0.032


```{r}
ldahist(bc_lda_org.pred$x, g=bc_d$Diagnosis)

bc_d$pred =bc_lda_org.pred$class
bc_d$miss = bc_lda_org.pred$class != bc_d$Diagnosis

sum(bc_d$miss)
```
#18개


##3d plot
```{r}
animate_xy(as.data.frame(bcX.sc), col=bc_d$pred,
           half_range=8)
```
#hard to distinguihs






####################################################################
#       T-sne
####################################################################
```{r}
library(liminal)
library(Rtsne)
library(umap)
tsne = Rtsne::Rtsne(bcX.sc,perplexity=80)
tsne_df = data.frame(tsneX=tsne$Y[,1],
                     tsneY=tsne$Y[,2])

plot(tsne_df, col=p_col, 
     pch=20, cex=1) + theme_minimal()

```


